
# Group K Project
From the data set, we plan to use columns sex, survived_200d, owner, growth, Std. BD, z.cones, and z.temp to identify the difference between male and female juveniles.

Hypothesis:
Male and female juvenile North American red squirrels Tamiasciurus hudsonicus differ
in their likelihood of acquiring a territory and surviving their first winter.

Predictions:
- One sex (male or female) might have a higher likelihood of acquiring a territory.
- Survival rates might be affected by sex due to differences in territorial acquisition, foraging, and resource allocation.

```{r}
library(tidyverse)
library(magrittr)
library(lmerTest)
library(MuMIn)
library(broom)
library(lme4)
```

# load data
```{r}
juveniles <- read_csv("C:/Users/Nemo/Desktop/EEB313/Group Project/squirrels dataset/autumn_juveniles.csv")

head(juveniles)
```
```{r}
autumn_juveniles <- juveniles %>% mutate(acquired_terr = case_when(owner == "acquired territory" ~ 1, TRUE ~ 0) )
head(autumn_juveniles)
```

```{r}
# number of male and females
n_sex <- autumn_juveniles %>%  group_by(sex) %>%
    tally()
n_sex
```

```{r}
# Doing a logistic regression for all populations
autumn_juveniles %>%
  group_by(sex) %>%
  ggplot(aes(x = acquired_terr, fill = sex)) + 
  geom_bar()
```


```{r}
model <- glm(acquired_terr ~ sex, 
             family = "binomial", 
             data = autumn_juveniles)
summary(model)
```

sexM:
Estimate = -0.39008
The negative value means males have lower odds of acquiring a territory compared to females.


### Study the effect of z.density to acquire_terr

### Latest comments and suggestions are going to be released this evening by Mete and Zoe










# Irrelevant wrong methods. Kept at the bottom
```{r}
#comparing between survival?
survive_autumn_juveniles <- autumn_juveniles %>% 
  filter(!is.na(survived_200d)) %>% 
  group_by(sex, survived_200d)

survive_autumn_juveniles
```

```{r}
# Calculate survival likelihood for each year and sex
survival_likelihood <- autumn_juveniles %>%
  group_by(year, sex) %>%
  summarise(likelihood = mean(survived_200d))

# Plot the survival likelihood
ggplot(survival_likelihood, aes(x = year, y = likelihood, color = sex)) +
  geom_line() +
  labs(title = "Long-term Survival Likelihood by Sex Over Years",
       x = "Year",
       y = "Survival Likelihood")
```

```{r}
# Calculate the likelihood of acquiring a habitat for each year and sex
acq_hab_ll <- autumn_juveniles %>%
  group_by(year, sex) %>%
  summarise(likelihood = mean(acquired_terr))

# Plot the survival likelihood
ggplot(acq_hab_ll, aes(x = year, y = likelihood, color = sex)) +
  geom_line() +
  labs(title = "Long-term Likelihood of acquiring a habitat by Sex Over Years",
       x = "Year",
       y = "Likelihood of Acquiring a habitat")
```



