---
title: "EEB313 project data self analysis"
output: pdf_document
date: "2024-11-19"
---

```{r}
library(tidyverse)
```


```{r}
data <- read.csv("autumn_juveniles.csv")
```


```{r}
data1 <- data %>%
  select(sex, survived_200d, owner, growth, Std.BD, z.cones, z.temp)
head(data1)
```
sex: The sex of the individual, typically 'M' for male or 'F' for female.
survived_200d: Indicates whether the individual survived 200 days (1 for survived, 0 for not survived).
owner: Status of territory ownership (e.g., "acquired territory" or "not yet acquired territory").
growth: Growth rate or a measure of growth for the individual.
Std.BD: Standardized body dimension or another standardized measurement related to the individual.
z.cones: Z-score of cone availability, related to food resources.
z.temp: Z-score of temperature, related to environmental conditions.

Prediction 1: One sex (male or female) might have a higher likelihood of acquiring a territory.

```{r}
data1 %>%
  group_by(sex, owner) %>%
  tally()
```

Female: 520/777 = 66.9%
Male: 363/628 = 57.8%
At first glance, female seems to have a higher likelihood of acquiring a territory.
But is it because of them being survived longer?

```{r}
data1 %>%
  group_by(sex, owner, survived_200d) %>%
  tally()
```

Does female or male lives longer?

```{r}
T3 <- data1 %>%
  group_by(sex, survived_200d)

table(T3$sex, T3$survived_200d)
chisq.test(T3$sex, T3$survived_200d)
```




Prediction 2: Survival rates might be affected by sex due to differences in territorial acquisition, foraging, and resource allocation.

```{r}

```


